Industrial, large-scale model predictive control with structured neural networks

### 2020-09-05 paper.tex
add a thesis.
some final changes for the first draft.

To gauge the performance of NNs in 
the validation simulations, and the offline
computational effort required to build the NNs, 
the following metrics are examined:

@@

\subsection{CSTRs with a flash separator}
For the first example, the two 
stirred tank reactors in series with a flash separator
depicted in Figure \ref{fig:schematic_cstrs} 
is considered as the plant.
In both the CSTRs, a desired product $B$ is produced 
with the reaction $A \rightarrow B$, and an undesired 
product $C$ is produced with the reaction $B \rightarrow C$. 
The reactant $A$ is the major component in the 
feed streams supplied to the 
reactors. After the second CSTR, the reaction mixture is 
sent to the non-adiabatic flash that separates the reactant $A$
from the product $B$. The $A$ rich vapor phase 
is recycled back to the first CSTR. 
The plant is simulated using the set of 
nonlinear ordinary differential equations (ODEs) available in 
\cite*{stewart:venkat:rawlings:wright:pannocchia:2010}.
The model has 12 states 
($H_r, x_{Ar}, x_{Br}, T_r, 
H_m, x_{Am}, x_{Bm}, T_m, H_b, x_{Ab}, x_{Bb}, T_b$),
6 control inputs 
($F_0, F_1, D, Q_r, Q_m, Q_b$), and 5 disturbances
($x_{A0}, x_{A1}, x_{B0}, x_{B1}, T_0$).
The controlled measurements with setpoints are the heights
and temperatures of the three units, and we
assume that all the states are measured for this example. 
The sample time for the measurements is chosen as 10 seconds. 
The parameter values used to simulate the ODEs for the plant, 
the actuator constraints, and the bounds of the controlled 
measurements and disturbances used for the data generation 
and validation simulations are 
shown in Table \ref{table:cstrs_pars}.

@@

A linear discrete time model with the chosen sample time
is obtained at the steady state shown in the table 
\ref{table:cstrs_pars}.
This model, inputs and outputs are scaled 
for the MPC controller such that
$\overline{u} - \underline{u} = \mathbf{2}$ and
$\overline{y} - \underline{y} = \mathbf{2}$.
The tuning parameters and 
control horizon length for the regulator are chosen as
$Q = 10^3 C'C$, $R = 0.1 I$, $S = 0.1 I$, 
and $N = 450$.
A total $1.5 \times 10^5$ samples of 
($x, u_{-1}, x_s, u_s, \kappa_N(\cdot)$) 
is generated for NN training, parallelized in 100 
separate offline simulations. The time consumed 
for this data generation step was 4.9 hours.

Four structured NN architectures 
are considered. Each NN
has three hidden layers with 
448 (NN-3-448), 480 (NN-3-480), 
512 (NN-3-512), and 544 (NN-3-544) 
nodes in each hidden 
layer respectively. For validation simulations, 
setpoint and disturbance signals 
with 24 setpoint changes and 48 disturbance changes 
is generated for 4320 timesteps (12 hours). 
The short horizon MPC controller is constructed with a 
horizon length of $N = 10$, and the architecture 
of the unstructured NN considered 
is three hidden layers with 224 nodes in each hidden layer. 
An additional scaling of $x$ and $x_s$ is performed for the 
NN training as $x := 2x/(x_{\textnormal{max}} - x_{\textnormal{min}})$ and 
$x_s := 2x_s/(x_{\textnormal{max}} - x_{\textnormal{min}})$,
in which $x_{\textnormal{max}}$ and $x_{\textnormal{min}}$
are the maximum and minimum values of the state 
observed in the training data.
This scaling is also performed in real-time 
when the NN is used to compute the actuator move.

@@

Second, the transient closed-loop 
performance of the trained NNs is examined
for one particular simulation. The Figures 
\ref{fig:cl_cstrs_inputs} and \ref{fig:cl_cstrs_outputs}
show the transient actuator 
and controlled outputs obtained with the NN-3-448 and 
optimal MPC controllers for a simulation period of 2 hours.
The Figure \ref{fig:cstrs_statistics} (center bottom)
shows the transient values of the controller performance index
obtained in this validation simulation
with the four structured NNs, SS, satK 
and optimal MPC controllers for the full 
simulation period of 12 hours. The closed-loop 
performance obtained using NNs is almost same as the optimal MPC 
controller and the degradation in the performance 
is quantified with the $\%$ performance loss metric.

Third, the online computational benefits of NNs 
over the QP based MPC is analyzed. The Figure 
\ref{fig:cstrs_statistics} (top right)
shows a histogram of online computation times
of the NNs and QP solver in the above validation
simulation. The average and worst-case speedups
obtained using NNs are summarized in the 
Table \ref{table:cstrs}. The QP solver takes 
about 9 to 14 seconds in this example and all the 
NNs compute the control in just 0.4 to 3 milliseconds, 
and are conveniently deployable in the 
sample time of 10 seconds.

The $\%$ performance loss 
obtained using the four heuristic controllers, 
their corresponding speedups, time required to 
train each NN, number of parameters in each NN, 
and the memory required to store  
the weights of each NN is summarized in the 
Table \ref{table:cstrs}. The $\%$ performance loss 
reported for the unstructured
NN is after training with $1.5 \times 10^5$  
samples and the performance is unacceptable 
for an industrial application. 
This noticeable performance improvement with the
structured NNs is because the control law 
is exactly the same as the optimal feedback law
at steady states, and the training process improves the 
approximation of the optimal feedback law
by the NN only for the transient states.

@@

The Figure \ref{fig:cdu_statistics} shows the 
statistics for the simulation study performed for this 
example. The data requirements to obtain 
an accurate NN controller is shown in top left.
The transient controller performance index 
for one validation simulation
obtained with the four structured NNs,
satK, and MPC controllers is shown in center bottom.
The top right plot shows the 
histogram of online computation times 
of the NNs and QP solver.
The Table \ref{table:cdu} gives a summary of 
the maximum time required to train the NNs, number of parameters
in each NN, memory footprint of the NNs, and 
the $\%$ performance loss metric obtained by 
the structured NNs, SS, satK, and SH 
controllers in the validation simulations.

The offline data generation and training times 
for the NNs shows that this large-scale example
is tractable with the proposed NN approach.
This large example is challenging, 
since on average the QP solver CVXOPT takes 37 seconds
and 57 seconds in the worst-case. In comparison,  
all the NNs execute MPC in about 2 to 6 milliseconds.
As observed from the plot of the transient controller 
performance index, the performance 
degradation by the NNs compared with 
the optimal MPC controller is negligible.
The sampling of the state-space for NN training
for one particular operational scenario 
with the corresponding setpoint and disturbance signals 
is the rule to obtain scalability of the NN controllers.

\section{Conclusion} \label{sec:conclusion}
Neural networks have been demonstrated 
to approximate the MPC feedback law 
for treating large-scale, linear MPC applications
that may be challeging with QP solvers. 
The proposed no online optimization-based 
methods in the literature such as storing 
the entire MPC feedback law 
and sampling the large dimensional state-space 
for the NN training are not tractable for the size of problems 
considered in this article.

The key features for scalability of the proposed approach 
are the structure of NN and data generation for only the 
typical plant operational scenarios. 
The NN design approach was demonstrated to scale on a
industrial crude distillation unit model with state and
control dimensions of 252 and 32 respectively, and 
a control-sample horizon length of 145 in the QP. 
Parallel computing was used for the data 
generation and graphical processing units for the NN 
training. After the offline design phase, NNs execute MPC 
about three to five orders of magnitude faster than an 
available QP solver 
with less than 1$\%$ performance degradation in the closed-loop. 
The NNs are easily deployable on memory constrained hardware.
Robustness properties
of NNs were established under bounded approximation errors of the optimal 
feedback law, that provides some theoretical support 
to practitioners interested in the deployment of NNs.

@@

%% New version of the num-names style
\bibliographystyle{elsarticle-num-names}
\bibliography{abbreviations,articles,books,unpub,proceedings}
\end{document}

>>>>

To gauge the performance of the NNs in 
the validation simulations, computational 
speedups over online QP based MPC, the offline
computational effort required to build the NNs,
and the memory required for the deployment of NNs,  
the following metrics are examined:

@@

\subsection{CSTRs with a flash separator}
For the first example, the two 
stirred tank reactors in series with a flash separator
depicted in Figure \ref{fig:schematic_cstrs} 
is considered as the plant.
In both the CSTRs, a desired product $B$ is produced 
with the reaction $A \rightarrow B$, and an undesired 
product $C$ is produced with the reaction $B \rightarrow C$. 
The reactant $A$ is the major component in the 
feed streams supplied to the 
reactors. After the second CSTR, the reaction mixture is 
sent to the non-adiabatic flash that separates the reactant $A$
from the product $B$. The $A$ rich vapor phase 
is recycled back to the first CSTR. 
The plant is simulated using the set of 
nonlinear ordinary differential equations (ODEs) available in 
\cite*{venkat:2006}, Appendix.
The model has 12 states 
($H_r, x_{Ar}, x_{Br}, T_r, 
H_m, x_{Am}, x_{Bm}, T_m, H_b, x_{Ab}, x_{Bb}, T_b$),
6 control inputs 
($F_0, F_1, D, Q_r, Q_m, Q_b$), and 5 disturbances
($x_{A0}, x_{A1}, x_{B0}, x_{B1}, T_0$).
The controlled measurements with setpoints are the heights
and temperatures of the three units, and we
assume that all the states are measured for this example. 
The sample time for the measurements is chosen as 10 seconds. 
The parameter values used to simulate the ODEs for the plant, 
the actuator constraints, and the bounds of the controlled 
measurements and disturbances used for the data generation 
and validation simulations are 
shown in Table \ref{table:cstrs_pars}.

@@

A linear discrete time model with the chosen sample time
is obtained at the steady state shown in the Table 
\ref{table:cstrs_pars}.
This model, inputs and outputs are scaled 
for the MPC controller such that
$\overline{u} - \underline{u} = \mathbf{2}$ and
$\overline{y} - \underline{y} = \mathbf{2}$.
The tuning parameters and 
control horizon length for the regulator are chosen as
$Q = 10^3 C'C$, $R = 0.1 I$, $S = 0.1 I$, 
and $N = 450$.
A total $1.5 \times 10^5$ samples of 
($x, u_{-1}, x_s, u_s, \kappa_N(\cdot)$) 
is generated for NN training, parallelized in 100 
separate offline simulations. The time consumed 
for this data generation step was 4.9 hours.

Four structured NN architectures 
are considered. Each NN
has three hidden layers with 
448 (NN-3-448), 480 (NN-3-480), 
512 (NN-3-512), and 544 (NN-3-544) 
nodes in each hidden 
layer respectively. For the validation simulations, 
setpoint and disturbance signals 
with 24 setpoint changes and 48 disturbance changes 
is generated for 4320 timesteps (12 hours). 
The short horizon MPC controller is constructed with a 
horizon length of $N = 10$, and the architecture 
of the unstructured NN considered 
is three hidden layers with 224 nodes in each hidden layer. 
An additional scaling of the 
states and target steady states ($x$ , $x_s$) is performed for the 
NN training as $x := 2x/(x_{\textnormal{max}} - x_{\textnormal{min}})$ and 
$x_s := 2x_s/(x_{\textnormal{max}} - x_{\textnormal{min}})$,
in which $x_{\textnormal{max}}$ and $x_{\textnormal{min}}$
are the maximum and minimum values of the state 
observed in the generated training data.
This scaling is also performed in real-time 
when the NN is used to compute the actuator move.

@@

Second, the transient closed-loop 
performance of the trained NNs is examined
for one particular simulation. The Figures 
\ref{fig:cl_cstrs_inputs} and \ref{fig:cl_cstrs_outputs}
show the transient actuator 
and controlled outputs obtained with the NN-3-448 and 
optimal MPC controllers for a simulation period of 2 hours.
The Figure \ref{fig:cstrs_statistics} (center bottom)
shows the transient values of the controller performance index
obtained in this validation simulation
with the four structured NNs, SS, satK 
and optimal MPC controllers for the full 
simulation period of 12 hours. The closed-loop 
performance obtained using NNs is almost same as the optimal MPC 
controller and the degradation in the performance 
is quantified with the $\%$ performance loss metric
shown in the Table \ref{table:cstrs}.

Third, the online computational benefits of NNs 
over the QP based MPC is analyzed. The Figure 
\ref{fig:cstrs_statistics} (top right)
shows a histogram of online computation times
of the NNs and the QP solver 
CVXOPT. The average and worst-case speedups
obtained using the NNs are summarized in the 
Table \ref{table:cstrs}. The QP solver takes 
about 9 to 14 seconds in this example and all the 
NNs compute the control in just 0.4 to 3 milliseconds, 
and are conveniently deployable in the 
sample time of 10 seconds.

The $\%$ performance loss 
obtained using the four fast controllers other 
than the structured NNs, 
their corresponding 
computational speedups, the maximum time required to 
train each NN, the number of parameters in each NN, 
and the memory required to store  
the weights of each NN is summarized in the 
Table \ref{table:cstrs}. The $\%$ performance loss 
reported for the unstructured
NN is after training with $1.5 \times 10^5$  
samples and the performance is unacceptable 
for an industrial application. 
This noticeable performance improvement with the
structured NNs is because the control law 
is exactly the same as the optimal feedback law
at the steady states, and that simplifies
the optimal feedback law approximation problem
since the training process only has to improve the 
NN control law for the transient states.

@@

The Figure \ref{fig:cdu_statistics} shows the 
statistics obtained in the simulation study performed for this 
example. The data requirements to obtain 
an accurate NN controller is shown in top left.
The transient controller performance index 
for one validation simulation
obtained with the four structured NNs,
satK, and MPC controllers is shown in center bottom.
The top right plot shows the 
histogram of online computation times 
of the NNs and the QP solver CVXOPT.
The Table \ref{table:cdu} gives a summary of 
the maximum time required to train the NNs, 
the number of parameters
in each NN, the memory footprint of the NNs, and 
the $\%$ performance loss obtained by 
the structured NNs, SS, satK, and SH 
controllers in the validation simulations.

The offline data generation and training times 
to build the NNs show that this large dimensional example
is tractable with the proposed NN approach.
This large example is challenging, 
since on average the QP solver CVXOPT takes 37 seconds
and 57 seconds in the worst-case. In comparison,  
all the NNs execute MPC in about 2 to 6 milliseconds.
As observed from the plot of the transient controller 
performance index, the performance 
degradation by the NNs relative to 
the optimal MPC controller is negligible.
The sampling of the state-space with 
the setpoint and disturbance signals
for one particular operational scenario,
combined with the offset-free NN structure
are key steps for scalability in this example.

\section{Conclusion} \label{sec:conclusion}
Neural networks have been demonstrated 
to approximate the MPC feedback law
for treating large-scale, linear MPC applications
that may be challeging with QP solvers. 
The proposed no online optimization-based
methods to-date in the literature such as storing 
the entire MPC feedback law 
and sampling the large dimensional state-space 
for the NN training are not tractable for the size of problems 
considered in this article.

The key features for the scalability of the proposed approach 
are the structure of NN and data generation for only the
typical plant operational scenarios. 
The NN design approach was demonstrated to scale on an
industrial crude distillation unit model with state and
control dimensions of 252 and 32 respectively, and 
a control-sample horizon length of 145 in the QP. 
After the offline design phase, NNs execute MPC 
about three to five orders of magnitude faster than an 
available QP solver 
with less than 1$\%$ performance degradation in the closed-loop. 
Parallel computing can be used for the data generation step
if solving QPs for each sampled state starts to take long 
amounts of time.
After the training process, 
the NNs are easily deployable on memory constrained hardware.
We established conditions under which the 
NN controllers are robust to state estimation errors 
and process disturbances,
that provides some theoretical support 
to practitioners interested in the deployment of NNs.

@@

%% New version of the num-names style
\bibliographystyle{elsarticle-num-names}
\bibliography{articles,books,unpub,proceedings}
\end{document}
